
```{r, echo = FALSE, include = FALSE}
library (tidyverse)
library (kableExtra)
library (downloadthis)

```


```{r, echo = FALSE, include = FALSE}
df2 <- rio::import ("data/df2.xlsx")
names (df2) <- c("subj",
                 "group",
                 "session",
                 "side",
                 "set",
                 "rep",
                 "atorq",
                 "awork",
                 "apow",
                 "ktorq",
                 "kwork",
                 "kpow")
```


# Exploration

## Descriptive statistics

You want to find some summary statistics like averages, standard deviation of some variables.

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj, group, session) %>%
  summarize (atorq = mean (atorq))

```


```{r, echo = FALSE, include = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/explore_data.xlsx")
```

### Syntax

Arguments to care about:

1. `VARIABLES`: Change the value `atorq` to the name of your dependent variable. You can even add many variables, with a space between them.

As the name suggests, I am asking SPSS to spit out the average, standard deviation, variance, range, minimum, and maximum values.

```{r, eval = FALSE}
DESCRIPTIVES VARIABLES=atorq
  /STATISTICS=MEAN STDDEV VARIANCE RANGE MIN MAX.


```

## QQ plot

You want to check for normality of distribution of a variable.

### Syntax

Arguments to care about:

1. `/VARIABLES`: Change the value `atorq` to the name of your dependent variable. You can add as many variables one after the other, each with a spacing.

```{r, eval = FALSE}
PPLOT
  /VARIABLES=atorq
  /NOLOG
  /NOSTANDARDIZE
  /TYPE=Q-Q
  /FRACTION=BLOM
  /TIES=MEAN
  /DIST=NORMAL.
```

