```{r, echo = FALSE, include = FALSE}
library (tidyverse)
library (kableExtra)
library (downloadthis)
```


```{r, echo = FALSE, include = FALSE}
df2 <- rio::import ("data/df2.xlsx")
names (df2) <- c("subj",
                 "group",
                 "session",
                 "side",
                 "set",
                 "rep",
                 "atorq",
                 "awork",
                 "apow",
                 "ktorq",
                 "kwork",
                 "kpow")
```


# Simpler statistics

## One-sample t test {#onettest}

**Hypothesis**  Is the ankle torque of my sample really different from 60 Nm/kg?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj) %>%
  summarize (atorq = mean (atorq))

```

```{r, eval  = FALSE , echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/one_sampleT_data.xlsx")
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

### Syntax

Arguments to care about:

1. `/VARIABLES`: Change the value `atorq` to the name of your dependent variable

2. `/TESTVAL`: Change the value to the mean you want to compare it against. In this example, I am asking *is my ankle torque of my sample really different from 0?*

```{r, eval = FALSE}
T-TEST
  /TESTVAL=60
  /MISSING=ANALYSIS
  /VARIABLES=atorq
  /ES DISPLAY(TRUE)
  /CRITERIA=CI(.95).
```

For reporting, please see Chapter \@ref(report-onettest).

### Non parametric equivalent

Arguments to care about:

1. Change the value `atorq` to the name of your dependent variable

```{r, eval = FALSE}
NPTESTS 
  /ONESAMPLE TEST (atorq) 
  /MISSING SCOPE=ANALYSIS USERMISSING=EXCLUDE
  /CRITERIA ALPHA=0.05 CILEVEL=95  SEED=RANDOM.
```


## Independent t-test{#twottest}

**Hypothesis**  Is the ankle torque from one group different from a separate group?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj, group) %>%
  summarize (atorq = mean (atorq)) 
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/two_sampleT_data.xlsx")
```

### Syntax

Arguments to care about:

1. `GROUPS`: Change the variable `group` to the name of your independent grouping variable. Also change `G` and `T` to the names of the levels within the grouping variable

2. `/VARIABLES`: Change `atorq` to the name of your dependent variable. 

```{r, eval = FALSE}
T-TEST GROUPS=group('G' 'T')
  /MISSING=ANALYSIS
  /VARIABLES=atorq
  /ES DISPLAY(TRUE)
  /CRITERIA=CI(.95).
```

For reporting, please see Chapter \@ref(report-twottest).


### Non parametric equivalent

Arguments to care about:

1. Change the value `atorq` to the name of your dependent variable, and `group` to the name of your independent variable.

```{r, eval = FALSE}
NPTESTS 
  /INDEPENDENT TEST (atorq) GROUP (group) 
  /MISSING SCOPE=ANALYSIS USERMISSING=EXCLUDE
  /CRITERIA ALPHA=0.05  CILEVEL=95.
```


## Paired t-test{#pairedttest}

**Hypothesis**  Is the ankle torque in the same group of subjects different after compared to before a training program?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  filter (group == "G") %>%
  group_by(subj, session) %>%
  summarize (atorq = mean (atorq)) %>% 
  pivot_wider(names_from = "session",
              values_from = "atorq")
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/paired_sampleT_data.xlsx")
```

### Syntax

Arguments to care about:

1. `PAIRS`: Change the variables `POST` and `PRE` to the name of the paired variables you are testing.  

```{r, eval = FALSE}
T-TEST PAIRS=POST WITH PRE (PAIRED)
  /ES DISPLAY(TRUE) STANDARDIZER(CORRECTED_SD)
  /CRITERIA=CI(.9500)
  /MISSING=ANALYSIS.
```

For reporting, please see Chapter \@ref(report-pairedttest).

### Non parametric equivalent

Arguments to care about:

1. Change the variables `POST` and `PRE` to the name of the paired variables you are testing. 

```{r, eval = FALSE}
NPTESTS 
  /RELATED TEST(POST PRE) 
  /MISSING SCOPE=ANALYSIS USERMISSING=EXCLUDE
  /CRITERIA ALPHA=0.05  CILEVEL=95.
```


## One-way between ANOVA{#oneBetwtest}

**Hypothesis**  Is the ankle torque different across the three independent groups? If yes, where do the differences lie? 

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  mutate (group = ifelse (group == "G" & session == "PRE",
                          1, ifelse (group == "G" & session == "POST",
                                     2, 3)))  %>%
  group_by(subj, group) %>%
  summarize (atorq = mean (atorq)) 
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/oneAnova_data.xlsx")
```

### Syntax

Arguments to care about:

1. Change the variables `atorq` and `group` to the names of your dependent and independent variables, respectively

```{r, eval = FALSE}
ONEWAY atorq BY group
  /ES=OVERALL
  /STATISTICS DESCRIPTIVES 
  /MISSING ANALYSIS
  /CRITERIA=CILEVEL(0.95)
  /POSTHOC=LSD T3 ALPHA(0.05).
```

For reporting, please see Chapter \@ref(report-oneBetwtest).

## One-way within ANOVA{#oneWithtest} 

**Hypothesis**  Is the ankle torque of the same group different across multiple sessions? If yes, where do the differences lie? 

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  mutate (session = ifelse (session == "PRE" & side == "R",
                          "pre", ifelse (session == "PRE" & side == "L",
                                     "post", "after"))) %>%
  group_by(subj, session) %>%
  summarize (atorq = mean (atorq)) %>%
  pivot_wider(names_from = c("session"),
              values_from = "atorq")
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/oneWithinAnova_data.xlsx")
```

### Syntax

Arguments to care about:

1. `GLM`. Add the names of all the variables (i.e. `pre`, `post`, `after`). These three variables actually represent the three categories of a common variable, which in this context we call `session`. You can given any name to this common variable.

2. `/WSFACTOR`. Change `session` to any name you want to denote the categories within your repeated measures.

3. `/MEASURE`. Change `atorq` to the name of your dependent variable. 

4. `/EMMEANS`. Change `session` to any name you want to denote the categories within your repeated measures.

5. `/WSDESIGN`. Change `session` to any name you want to denote the categories within your repeated measures.

```{r, eval = FALSE}
GLM pre post after
  /WSFACTOR=session 3 Polynomial 
  /MEASURE=atorq 
  /METHOD=SSTYPE(3)
  /EMMEANS=TABLES(session) COMPARE ADJ(LSD)
  /PRINT=ETASQ PARAMETER HOMOGENEITY 
  /CRITERIA=ALPHA(.05)
  /WSDESIGN=session.
```

For reporting, please see Chapter \@ref(report-oneWithtest).

## Correlation{#corr} 

**Hypothesis** Are the variables of ankle torque, ankle work, knee torque, and knee work correlated with each other? 

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  slice (1:50) %>%
  select (subj, atorq:kpow) %>%
  mutate (subj = 1:50)
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/corr_data.xlsx")
```

### Syntax

Arguments to care about:

1. `/VARIABLE`. Add the names of all the variables (i.e. `atorq`, `awork`, `apow`, `ktorq`) you want to find a correlation. 

```{r, eval = FALSE}
CORRELATIONS
  /VARIABLES=atorq awork apow ktorq
  /PRINT=TWOTAIL NOSIG FULL
  /CI CILEVEL(95) BIAS(TRUE)
  /MISSING=PAIRWISE.

```


For reporting, please see Chapter \@ref(report-corr).