
```{r, echo = FALSE, include = FALSE}
library (tidyverse)
library (kableExtra)
library (downloadthis)
```


```{r, echo = FALSE, include = FALSE}
df2 <- rio::import ("data/df2.xlsx")
names (df2) <- c("subj",
                 "group",
                 "session",
                 "side",
                 "set",
                 "rep",
                 "atorq",
                 "awork",
                 "apow",
                 "ktorq",
                 "kwork",
                 "kpow")
```


# Tougher statistics

## Two (or more) way between ANOVA{#twoBetwtest}

**Hypothesis**  Do different strength training methods (`G` and `T`) result in different ankle torque, and is the difference dependent on supplement intake (`NO` and `YES`)?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  mutate (supplement = ifelse (session == "PRE", "NO", "YES")) %>%
  group_by(subj, group, supplement) %>%
  summarize (atorq = mean (atorq)) 
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/2wayBetweenAnova_data.xlsx")
```

### Syntax

Arguments to care about:

1. Replace all occurrences of `atorq` with the name of your dependent variable.

2. Replace all occurrences of `group` and `supplement` with the names of your independent variables, respectively. 

```{r, eval = FALSE}
UNIANOVA atorq BY group supplement
  /METHOD=SSTYPE(3)
  /INTERCEPT=INCLUDE
  /EMMEANS=TABLES(group) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(supplement) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(group*supplement) COMPARE(group) ADJ(LSD)
  /EMMEANS=TABLES(group*supplement) COMPARE(supplement) ADJ(LSD)
  /PRINT ETASQ PARAMETER HOMOGENEITY
  /CRITERIA=ALPHA(.05)
  /DESIGN=group supplement group*supplement.
```

For reporting, please see Chapter \@ref(report-twoBetwtest).

## Two (or more) way within ANOVA{#twoWithtest} 

**Hypothesis**  Does strength training result in an improvement over a training session (`PRE`, `POST`) in the same group of people, and is the improvement dependent upon side (`L`, `R`)? 

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj, session, side) %>%
  summarize (atorq = mean (atorq)) %>%
  pivot_wider(names_from = c("session", "side"),
              values_from = "atorq")
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/twoWithinAnova_data.xlsx")
```

### Syntax

Arguments to care about:

1. Notice I have four names: `POST_L`, `POST_R`, ` PRE_L`, `PRE_R`. See the pattern? It is actually made up of two variables which we call here `session` (`PRE`, `POST`) and `side` (`L`, `R`). Change the column names to the merged names of all the independent within-subject variables you have. 

2. Replace `session` and `side` to any given name to these two common variables.


```{r, eval = FALSE}
GLM POST_L POST_R PRE_L PRE_R
  /WSFACTOR=session 2 Polynomial side 2 Polynomial 
  /METHOD=SSTYPE(3)
  /EMMEANS=TABLES(session) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(side) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(session*side) COMPARE(session) ADJ(LSD)
  /EMMEANS=TABLES(session*side) COMPARE(side) ADJ(LSD)
  /PRINT=ETASQ PARAMETER HOMOGENEITY 
  /CRITERIA=ALPHA(.05)
  /WSDESIGN=session side session*side.
```

For reporting, please see Chapter \@ref(report-twoWithtest).

## Repeated (within-between) measures ANOVA{#RManovatest} 

**Hypothesis** Do different strength training methods (`G` and `T`) result in different ankle torque changes over session (`PRE`, `POST`)?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj, group, session) %>%
  summarize (atorq = mean (atorq)) %>%
  pivot_wider(names_from = "session",
              values_from = "atorq")
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/rmAnova_data.xlsx")
```

### Syntax

Arguments to care about:

1. Replace all occurrences of `atorq` with the name of your dependent variable.

2. Replace all occurrences of `group` with the name of your between group independent variable. 

3. Replace  `PRE` and `POST` with the names of your columns which contains the repeat measured variable. You can have more than two really.

4. All the repeated measured variable actually reflect categories within a single variable, which in this context we call `session`. You can given any name to this common variable.

5. Change this common name, in this case `session`, to the name you have chosen in (4).

```{r, eval = FALSE}
GLM PRE POST BY group
  /WSFACTOR=session 2 Polynomial 
  /MEASURE=atorq 
  /METHOD=SSTYPE(3)
  /POSTHOC=group(LSD T3) 
  /PLOT=PROFILE(group*session) TYPE=LINE ERRORBAR=CI MEANREFERENCE=NO YAXIS=AUTO
  /EMMEANS=TABLES(group) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(session) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(group*session) COMPARE (session)ADJ(LSD)
  /EMMEANS=TABLES(group*session) COMPARE (group)ADJ(LSD)
  /PRINT=DESCRIPTIVE ETASQ PARAMETER HOMOGENEITY 
  /CRITERIA=ALPHA(.05)
  /WSDESIGN=session 
  /DESIGN=group.
```

For reporting, please see Chapter \@ref(report-RManovatest).

## Linear mixed model{#lmmtest} 

**Hypothesis** Do different strength training methods (`G` and `T`) result in different ankle torque changes over session (`PRE`, `POST`)?

### What your data needs to look like?

```{r, echo = FALSE, include = FALSE}
df_sub <- df2 %>%
  group_by(subj, group, session) %>%
  summarize (atorq = mean (atorq)) 
```

```{r, echo = FALSE}
df_sub %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

df_sub %>%
  download_this(
    output_name = "data",
    output_extension = ".xlsx",
    button_label = "Download data",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save"
  )
```

```{r, eval  = FALSE, echo = FALSE, include = FALSE}
writexl::write_xlsx(df_sub, 
                    "data/mixedModel_data.xlsx")
```

### Syntax

Arguments to care about:

1. Replace all occurrences of `atorq` with the name of your dependent variable.

2. Replace all occurrences of `group` and `session` with the names of your independent variables, respectively. 

3. Replace all occurrences of `subj` with the name of your identifier variable - i.e. the name of the column that contains the unique patient name or number.

```{r, eval = FALSE}
MIXED atorq BY group session
  /CRITERIA=DFMETHOD(SATTERTHWAITE) CIN(95) MXITER(100) MXSTEP(10) SCORING(1) 
    SINGULAR(0.000000000001) HCONVERGE(0.00000001, RELATIVE) LCONVERGE(0, ABSOLUTE) PCONVERGE(0, 
    ABSOLUTE)
  /FIXED=group session group*session | SSTYPE(3)
  /METHOD=REML
  /PRINT=DESCRIPTIVES  SOLUTION
  /RANDOM=INTERCEPT | SUBJECT(subj) COVTYPE(VC) SOLUTION
  /REPEATED=session | SUBJECT(subj) COVTYPE(DIAG)
  /EMMEANS=TABLES(group) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(session) COMPARE ADJ(LSD)
  /EMMEANS=TABLES(group*session) COMPARE(group) ADJ(LSD)
  /EMMEANS=TABLES(group*session) COMPARE(session) ADJ(LSD).
```

For reporting, please see Chapter \@ref(report-lmmtest).